---
title: "Midterm"
author: "Nila Nathan"
date: "10/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages('foreign')
#library(foreign)
#DIQdata <- read.xport("V:/DIQ_I (1).XPT")
```

```{r}
#install.packages('SASxport')
#install.packages('Hmisc')
library(SASxport)
DIQdata <-read.xport("V:/DIQ_I (1).XPT")
View(DIQdata)
```


```{r}
#fixing column DID040

library(tidyverse)
library(dplyr)
library(magrittr)
DIQdata$DID040[is.na(DIQdata$DID040)] = median(DIQdata$DID040, na.rm = TRUE)
DIQdata %<>% mutate(DID040 = ifelse(DIQ010 == 2 , "Not Applicable", 
                                    ifelse(DIQ010 == 3 , "Not Applicable",
                                           ifelse(DIQ010 == 7 , "Not Applicable",
                                                  ifelse(DIQ010 == 9 , "Not Applicable", DID040)))))

```


```{r}
#fixing column DIQ160

DIQdata$DIQ160[is.na(DIQdata$DIQ160)] = median(DIQdata$DIQ160, na.rm = TRUE)

DIQdata %<>% mutate(DIQ160 = ifelse(DIQ010 ==3 , 1, DIQ160))
DIQdata %<>% mutate(DIQ160 = ifelse(DID040 <12 , "Not Applicable", DIQ160))
DIQdata %<>% mutate(DIQ160 = ifelse(DIQ010==1 , "Not Applicable", DIQ160))
DIQdata %<>% mutate(DIQ160 = ifelse(DID040 >=12 & DIQ010==3, "Not Applicable", DIQ160))


```
For column DIQ160 I first performed a median imputation. Then, I filled in cells by putting anyone who previously put down 'borderline' down also as 'prediabetes'. Other restrictions simply came from the DIQ-check item portion

```{r}
#fixing column DIQ170 and DIQ172

#median(DIQdata$DIQ170, na.rm = TRUE) - wanted to make sure the median was 2 before imputation
#median(DIQdata$DIQ172, na.rm = TRUE) - wanted to make sure the median was 2 before imputation


DIQdata$DIQ170[is.na(DIQdata$DIQ170)] = median(DIQdata$DIQ170, na.rm = TRUE)
DIQdata$DIQ172[is.na(DIQdata$DIQ172)] = median(DIQdata$DIQ172, na.rm = TRUE)

```

Decided to use median imputations to fill in those 

```{r}
#fixing columns DIQ175

dat = data.frame(A=DIQdata$DIQ175A, B=DIQdata$DIQ175B, C=DIQdata$DIQ175C, D=DIQdata$DIQ175D, E=DIQdata$DIQ175E, f=DIQdata$DIQ175F, G=DIQdata$DIQ175G, H=DIQdata$DIQ175H, I=DIQdata$DIQ175I, J=DIQdata$DIQ175J, K=DIQdata$DIQ175K, L=DIQdata$DIQ175L, M=DIQdata$DIQ175M, N=DIQdata$DIQ175N, O=DIQdata$DIQ175O, P=DIQdata$DIQ175P,Q=DIQdata$DIQ175Q, R=DIQdata$DIQ175R, S=DIQdata$DIQ175S, t=DIQdata$DIQ175T, U=DIQdata$DIQ175U, V=DIQdata$DIQ175V, W=DIQdata$DIQ175W, X=DIQdata$DIQ175X)

DIQdata$DIQ175 <- paste(dat$A, dat$B, dat$C, dat$D, dat$E, dat$f, dat$G, dat$H, dat$I, dat$J, dat$K, dat$L, dat$M, dat$N, dat$O, dat$P, dat$Q, dat$R, dat$S, dat$t, dat$U, dat$V, dat$W, dat$X, sep="_", na.rm=TRUE)



DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("_NA", "",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("_TRUE", "",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("NA_","",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("_",",",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("10","Family History",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("11","Overweight",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("12","Age",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("13","Poor Diet",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("14","Race",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("15","Had a baby weighed over 9lbs over birth",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("16","Lack of physical activity",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("17","High blood pressure",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("18","High blood sugar",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("19","High cholesterol",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("20","Hypglocemic",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("21","Extreme Hunger",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("22","Tingling/Numbness in hands or feet",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("23","Blurred vision",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("24","Increased fatigue",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("25","Anyone could be at risk",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("26","Doctor warning",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("27","Other, specify",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("28","Gestational Diabetes",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("29","Frequent Urination",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("30","Thirst",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("31","Craving for sweet/eating a lot of sugar",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("32","Medication",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("33","Polycystic Ovarian Syndrome",x))
DIQdata$DIQ175=sapply(DIQdata$DIQ175,function(x)gsub("NA","Not Applicable",x))

#install.packages("labelled")
library(labelled)

var_label(DIQdata$DIQ175) <- "Reason for feeling at risk/being told at risk"

DIQdata$DIQ175A=NULL
DIQdata$DIQ175B=NULL
DIQdata$DIQ175C=NULL
DIQdata$DIQ175D=NULL
DIQdata$DIQ175E=NULL
DIQdata$DIQ175F=NULL
DIQdata$DIQ175G=NULL
DIQdata$DIQ175H=NULL
DIQdata$DIQ175I=NULL
DIQdata$DIQ175J=NULL
DIQdata$DIQ175K=NULL
DIQdata$DIQ175L=NULL
DIQdata$DIQ175M=NULL
DIQdata$DIQ175N=NULL
DIQdata$DIQ175O=NULL
DIQdata$DIQ175P=NULL
DIQdata$DIQ175Q=NULL
DIQdata$DIQ175R=NULL
DIQdata$DIQ175S=NULL
DIQdata$DIQ175T=NULL
DIQdata$DIQ175U=NULL
DIQdata$DIQ175V=NULL
DIQdata$DIQ175W=NULL
DIQdata$DIQ175X=NULL


```

Here I combined all the columns of 175A-175X and put into one column titled 175 and instead of the numbers, made it more clear but substituting what the numbers meant

```{r}
#fixing column DID060 and DIQ060U

dat6 = data.frame(insulin=DIQdata$DID060, leng=DIQdata$DIQ060U)
DIQdata$DIQ060 = paste(dat6$insulin, dat6$leng, sep="/", na.rm=TRUE)
DIQdata$DIQ060=sapply(DIQdata$DIQ060,function(x)gsub("/TRUE", "",x))
DIQdata$DIQ060=sapply(DIQdata$DIQ060,function(x)gsub("NA/NA", NA,x))
DIQdata$DIQ060=sapply(DIQdata$DIQ060,function(x)gsub("0/NA", "Not Applicable",x))
DIQdata$DIQ060=sapply(DIQdata$DIQ060,function(x)gsub("/1", "/day",x))
DIQdata$DIQ060=sapply(DIQdata$DIQ060,function(x)gsub("/2", "/week",x))
DIQdata$DIQ060=sapply(DIQdata$DIQ060,function(x)gsub("/3", "/month",x))
DIQdata$DIQ060=sapply(DIQdata$DIQ060,function(x)gsub("/4", "/year",x))



DIQdata %<>% mutate(DIQ060 = ifelse(DIQ050 == 2 , "Not Applicable", 
                                    ifelse(DIQ050 == 7 , "Not Applicable",
                                           ifelse(DIQ050 == 9 , "Not Applicable", DIQ060))))

DIQdata$DID060=NULL
DIQdata$DIQ060U=NULL

```

Also combined column of insulin and units. This got rid of a few missing values.

```{r}
#fixing column DIQ070

DIQdata$DIQ070[is.na(DIQdata$DIQ070)] = median(DIQdata$DIQ070, na.rm = TRUE)

DIQdata %<>% mutate(DIQ070 = ifelse(DIQ010 == 2 , "Not Applicable", 
                                    ifelse(DIQ010 ==7 , "Not Applicable",
                                           ifelse(DIQ010 == 9 , "Not Applicable",
                                              ifelse(DIQ160 == 2 , "Not Applicable", 
                                                     ifelse(DIQ160 == 7 , "Not Applicable",
                                                            ifelse(DIQ160 == 9 , "Not Applicable",DIQ070)))))))

```




```{r}
#fixing column 230

DIQdata %<>% mutate(DIQ230 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DIQ230)))


```

```{r}
#fixing column 240

DIQdata %<>% mutate(DIQ240 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DIQ240)))

```

```{r}
#fixing column 250

DIQdata %<>% mutate(DID250 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DID250)))

DIQdata %<>% mutate(DID250 = ifelse(DIQ240 == 2 , "Not Applicable", 
                                    ifelse(DIQ240 == 7 , "Not Applicable", 
                                           ifelse(DIQ240==9, "Not Applicable",DID250))))

```

```{r}
#fixing column 260

dat4 = data.frame(G=DIQdata$DID260, U=DIQdata$DIQ260U)
DIQdata$DIQ260 <- paste(dat4$G, dat4$U, sep="/", na.rm=TRUE)
DIQdata$DIQ260=sapply(DIQdata$DIQ260,function(x)gsub("/TRUE", "",x))
DIQdata$DIQ260=sapply(DIQdata$DIQ260,function(x)gsub("NA/NA", NA,x))
DIQdata$DIQ260=sapply(DIQdata$DIQ260,function(x)gsub("0/NA", "Not Applicable",x))
DIQdata$DIQ260=sapply(DIQdata$DIQ260,function(x)gsub("/1", "/day",x))
DIQdata$DIQ260=sapply(DIQdata$DIQ260,function(x)gsub("/2", "/week",x))
DIQdata$DIQ260=sapply(DIQdata$DIQ260,function(x)gsub("/3", "/month",x))
DIQdata$DIQ260=sapply(DIQdata$DIQ260,function(x)gsub("/4", "/year",x))

DIQdata %<>% mutate(DIQ260 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DIQ260)))

DIQdata %<>% mutate(DIQ260 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DIQ260)))

DIQdata$DID260=NULL
DIQdata$DIQ260U=NULL

```

Combined columns like before in order to make it look cleaner (and when it joins together it also gets rid of some missing values.)


```{r}
#fixing column 275

DIQdata %<>% mutate(DIQ275 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DIQ275)))
```

```{r}
#fixing column DIQ280

DIQdata %<>% mutate(DIQ280 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DIQ280)))

DIQdata %<>% mutate(DIQ280 = ifelse(DIQ275 == 2 , "Not Applicable", 
                                    ifelse(DIQ275 == 7 , "Not Applicable", 
                                           ifelse(DIQ275==9, "Not Applicable",DIQ280))))
```


```{r}
#fixing column DIQ291

DIQdata %<>% mutate(DIQ291 = ifelse(DIQ010 == 3 , "Not Applicable", 
                                    ifelse(DIQ160 == 1 , "Not Applicable", DIQ291)))

DIQdata %<>% mutate(DIQ291 = ifelse(DIQ275 == 2 , "Not Applicable", 
                                    ifelse(DIQ275 == 7 , "Not Applicable", 
                                           ifelse(DIQ275==9, "Not Applicable",DIQ291))))
```


```{r}
#fixing columns 300

DIQdata$DIQ300S[is.na(DIQdata$DIQ300S)] = round(mean(DIQdata$DIQ300S[which(DIQdata$DIQ300S!=9999)]))
DIQdata$DIQ300D[is.na(DIQdata$DIQ300D)] = round(mean(DIQdata$DIQ300D[which(DIQdata$DIQ300D!=9999)]))

dat2 = data.frame(SBP=DIQdata$DIQ300S, DBP=DIQdata$DIQ300D)
DIQdata$DIQ300 <- paste(dat2$SBP, dat2$DBP, sep="/", na.rm=TRUE)
DIQdata$DIQ300=sapply(DIQdata$DIQ300,function(x)gsub("/TRUE", "",x))
DIQdata$DIQ300=sapply(DIQdata$DIQ300,function(x)gsub("NA/NA", NA,x))

DIQdata %<>% mutate(DIQ300 = ifelse(DID040 < 12 , "Not Applicable", DIQ300))

DIQdata$DIQ300S=NULL
DIQdata$DIQ300D=NULL


```

Did an imputation first because I assumed people who were taking diabetes survey probably had their blood pressure checked at some point. Combined columns like before in order to make it look cleaner (and when it joins together it also gets rid of some missing values.)


```{r}
#fixing columns 310

DIQdata$DID310S[is.na(DIQdata$DID310S)] = round(mean(DIQdata$DID310S[which(DIQdata$DID310S<6666)]))
DIQdata$DID310D[is.na(DIQdata$DID310D)] = round(mean(DIQdata$DID310D[which(DIQdata$DID310D<6666)]))

dat3 = data.frame(SBP2=DIQdata$DID310S, DBP2=DIQdata$DID310D)
DIQdata$DID310 <- paste(dat3$SBP2, dat3$DBP2, sep="/", na.rm=TRUE)
DIQdata$DID310=sapply(DIQdata$DID310,function(x)gsub("/TRUE", "",x))
DIQdata$DID310=sapply(DIQdata$DID310,function(x)gsub("NA/NA", NA,x))

DIQdata %<>% mutate(DID310 = ifelse(DID040 < 12 , "Not Applicable", DID310))


DIQdata$DID310S=NULL
DIQdata$DID310D=NULL

```

Same reasoning as above.

```{r}
#fixing column 320

DIQdata %<>% mutate(DID320 = ifelse(DID040 < 12 , "Not Applicable", DID320))


```

```{r}
#fixing column 330

DIQdata %<>% mutate(DID330 = ifelse(DID040 < 12 , "Not Applicable", DID330))

DIQdata %<>% mutate(DID330 = ifelse(DID320 == 5555 , "Not Applicable", 
                                    ifelse(DID320 == 6666 , "Not Applicable", DID330)))
```

```{r}
#fixing column 341

DIQdata %<>% mutate(DID341 = ifelse(DID040 < 12 , "Not Applicable", DID341))

```


```{r}
#fixing column 350

dat5 = data.frame(feet=DIQdata$DID350, times=DIQdata$DIQ350U)
DIQdata$DIQ350 <- paste(dat5$feet, dat5$times, sep="/", na.rm=TRUE)
DIQdata$DIQ350=sapply(DIQdata$DIQ350,function(x)gsub("/TRUE", "",x))
DIQdata$DIQ350=sapply(DIQdata$DIQ350,function(x)gsub("NA/NA", NA,x))
DIQdata$DIQ350=sapply(DIQdata$DIQ350,function(x)gsub("0/NA", "Not Applicable",x))
DIQdata$DIQ350=sapply(DIQdata$DIQ350,function(x)gsub("/1", "/day",x))
DIQdata$DIQ350=sapply(DIQdata$DIQ350,function(x)gsub("/2", "/week",x))
DIQdata$DIQ350=sapply(DIQdata$DIQ350,function(x)gsub("/3", "/month",x))
DIQdata$DIQ350=sapply(DIQdata$DIQ350,function(x)gsub("/4", "/year",x))

DIQdata %<>% mutate(DIQ350 = ifelse(DID040 < 12 , "Not Applicable", DIQ350))

DIQdata$DID350=NULL
DIQdata$DIQ350U=NULL


```

Once again, combined columns to get rid of missing values and make it cleaner.


```{r}
#fixing column 360

DIQdata %<>% mutate(DIQ360 = ifelse(DID040 < 12 , "Not Applicable", DIQ360))

```


```{r}
#fixing column 080

DIQdata %<>% mutate(DIQ080 = ifelse(DID040 < 12 , "Not Applicable", DIQ080))

```


```{r}

write.csv(DIQdata, "V:/DIQdata.csv")

```

Here I saved it to import into SQL where I can rename the columns. 

```{r}
#Connecting to SQL database

library('RODBC')
myconn=RODBC::odbcConnect('QBS181','nnathan','nnathan@qbs181')
library(sqldf)
library(dbplyr)
```

```{r}
#Retrieving data from SQL with renamed columns

R_DIQ=sqlQuery(myconn, "select * from nnathan.DIQdata")
head(R_DIQ,10)
```

